### 2020-12-10 slides.tex

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph:
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={insert title here},
  pdfborder={0 0 0},
  breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\newif\ifbibliography
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-2}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{insert title here}
\date{}

\begin{document}
\frame{\titlepage}

\begin{frame}{NHL Background}
\protect\hypertarget{nhl-background}{}

\begin{block}{The National Hockey League (NHL)}

\begin{itemize}
\tightlist
\item
  Each team plays 82 games per season
\item
  This allows for every team to play each team at least twice
\end{itemize}

\end{block}

\begin{block}{The changes for the 2013-14 season}

\begin{itemize}
\tightlist
\item
  Winnipeg Jets move from the Eastern to the Western confrence
\item
  Columbus Blue jackets and Detroit Redwings move from the Wet to the
  East\\
\end{itemize}

\end{block}

\begin{block}{The reasoning for the sampling}

\begin{itemize}
\tightlist
\item
  Changes left the West with 2 divisions of 14 teams and the East with 2
  divisions of 16 teams
\item
  8 teams from both confrences wer garunteed to make the playoffs
\item
  The sampling is meant to test wheter or not it is more difficult to
  reach the playoffs as an eastern confrence team
\end{itemize}

\end{block}

\end{frame}

\begin{frame}{Extension 2017-18 season}
\protect\hypertarget{extension-2017-18-season}{}

\begin{itemize}
\tightlist
\item
  This season added the Vegas Golden Knights
\item
  Now the west has 1 divion of 8 teams and one division of 7 teams
\item
  This creates new probabillities for the required points to make it to
  the playoffs
\end{itemize}

\end{frame}

\end{document}