### 2020-12-10 slides.Rmd
added more slides

---
title: "Using Monte Carlo to estimate the effects of NHL rule changes"
author: "Kilbourne Charrier, Michael Bowers, Clinton, Sam"
output: slidy_presentation 
---

## The National Hockey League (NHL)

- Each team plays 82 games per season
- This allows for every team to play each team at least twice
- The league is split up into two conferences, the East and the West, each with two divisions
- The East has the Metropolitan and the Atlantic divisions while the West has the Central and Pacific divisions.

## The Changes for the 2013-14 season

- Winnipeg Jets move from the Eastern to the Western confrence
- Columbus Blue jackets and Detroit Redwings move from the West to the East
- Each conference previoulsy had three divisions of five teams but these were condensed into the four divisions that currently exist

# Playoff Structure
- For the previous conference/division structure, the top two teams in each division secure a berth into the playoffs
- The remaining four slots are claimed by the two teams in each conference with the next highest point total regardles of the division. These are known as the wild cards.
- The process is similar with the new divisions in each conference
- Now the top three teams in each division secure a playoff berth and the wild cards are chosen using the same process as before

# The Reasoning for the Sampling

- Changes left the West with two divisions of 14 teams and the East with 2 divisions of 16 teams
- 8 teams from both conferences were garunteed to make the playoffs
- The sampling is meant to test whether or not it is more difficult to reach the playoffs as an eastern confrence team

# The Model: Goals

- To simulate the performance of each team over a season with varying rules
- To create an accurate scoring system that accounts for ties

# The Model: Goals

- To simulate the performance of each team over a season with varying rules
- To create an accurate scoring system that accounts for ties
- (For me) to learn how hockey works!

# The Model: Methods

- We generate each team's "quality" from a normal distribution
- For each of the games played we generate a "game performance" metric from a normal distribution with the mean being the respective team's quality.

![](Images/teamQuality.png)
![](Images/teamPerformance.png)

# The Model: Methods

- A game will go into overtime ~ 22.4% of the time
- If the teams skills are within a threshold alpha, it will go into overtime
- Once in overtime rescale performances and determine winner by weighted coinflip

# The Model: Methods

- We simulate 10,000 seasons to gather our metrics
- Sort teams into the right conference, then into the right division based on which set of rules we are looking at (old or new)
- Calculate the conference gap

# 2013-2014 Season Results

- We have found a noticably large conference gap, with the mean of the old alignment being about -0.32 and the mean of the new alignment being about 2.18.

![](Images/GapVis.png)

# 2013-2014 Season Results

- Another interesting graph!

![](Images/SeedPlace.png)

# Extension 2017-18 season

- This season added the Vegas Golden Knights
- Now the West has one divion of 8 teams and one division of 7 teams
- This creates new probabilities for the required points to make it to the playoffs