### 2020-12-14 Paper.Rmd
Update Paper.Rmd

---
title: "NHL realignment effects on conference gap"
author: "Kilbourne Charrier, Michael Bauers, Clinton Rothschild, Sam Downs"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Motivation
   In the 2013-14 National Hockey League(NHL) season, the league made some changes to the confrence numbers. The Winnipeg Jets moved from the Eastern confrence to the Western confrence. This createed an imbalance in the confrence numbers giving the Eastern confrence  teams and leaving the Western confrence with only fourteen teams. From the start, it already seemed unfair to have more teams in one confrence than the other. If every team had equal skill, the fourteen teams in the western confrence had an equal one four in seven chance of making it to the playoffs. In contrast, the sixteen teams in the eastern confrence had a fifty percent chance of making it to the playoffs. This uneven probabillity will be known for the rest of the paper as the confrence gap.  
  Now how could an organization as big as the NHL allow for that kind of discrepency between the confrences? That is what had us curious. As a continuation of our research we wanted to look at a nother season with a similiar balance shift in the confrences to see how that change couild have affected the outcome of the season. The season wwe choose was the recent expansion draft season of 2017-18. Before the start of the season, the NHL added a new franchis in the Las Vegas Golden Knights. This tewam was set to join the Western Confrence. This shifted the confrence split from fourteen to sixtten West to East to a closer fifteen to sixteen split. We wanted to see if our result was statistically significant from both the confrence gap of an even season and the confrence gap from our tested 2013-14 season.

## Introduction
  In 2011 when the Atlanta Thrashers became the Winnipeg Jets, a problem was presented as Winnipeg is considerably far from the east coast of the North american continent. Because of this and how the season is set up, the average travel time for teams playing winnipeg were fairly large as eastern confrence teams play echother more often. Just before the start of the 2013-14 season, the league offices for the NHL decided to move over the Winnipeg Jets from the Eastern Confrence to the Western Confrence and in return the Eastern Confrence aquired the Columbus Blue Jackets and the Detroit Redwings. This left the the two confrences at an uneven balance. The Eastern Confrence now had sixteen teams while the Western Confrence only had fourteen teams. This presented a curious problem as for both confrences, eight teams make the playoff brackets. It is easy to see the problem at hand that the western confrence has just as many playoff slots for having two less teams. The goal of our research and the paer we leaned on, was to discover what that difference could mean in the grander sense of the league moving forward with this set up. We did this based on the average points and wins necessary to make the playoffs for both confrences.
  To further our study, we also looked at the 2017-18 season. This season is important becasue it gave the introduction of a new team in the league none as the Las Vegas Golden Knights. This team was added to the Western Confrence makiing the split slightly more even. Our goal for this was to see how adding a single team could change the skill outcome for the playoffs. We looked at the same measures to understand the new confrence gap as our previous analysis of the 2013-14 season.

# Methodology

# Results